pirouette: Determining the error BEAST2 makes in inferring a phylogeny

### 2019-03-12 article.tex
Updates from Overleaf

\begin{sidewaystable}
\centering
  \begin{tabular}{|@{}c|p{2.5cm}|p{9cm}|p{4.5cm}@{}|}
    \hline
    \centering
    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} & \textbf{Possible values} \\ 
    \hline
    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model & Yule, BD, CBS, CCP, CEP \\
    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates & strict, relaxed log-normal \\
    $\mathit{s_{m}}$ & site model & Nucleotide substitution model & JC, HKY, TN93, GTR \\
    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs & $m_{r} \in \mathbb{R}_{>0}$\\
    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree & any combination of a, c, g, t \\
    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common ancestor & see documentation \\
    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup & see documentation \\
    $\mathit{m_{t}}$ & model type & Criterion to select an inference model & Generative, Candidate \\
    $\mathit{r_{i}}$ & run if & Condition under which an inference model is used & Always, Best candidate\\
    $\mathit{e_{f}}$ & error function & Specifies how to measure the error & nLTT, $|\gamma|$ \\
    $\mathit{t_{m}}$ & twin model & Sets the tree prior for the twin tree & Yule, BD \\
    \hline
  \end{tabular}
  \caption{
    Most important parameter options. BD = birth death, CBS = coalescent
    Bayesian Skyline, CCP = coalescent constant-population, CEP = coalescent
    exponential-population, JC = Jukes-Cantor 1969 (\cite{jukes1969evolution}), HKY = Hasegawa, Kishino and 
    Yano (\cite{hasegawa1985dating}), TN93 = \cite{tamura1993estimation}, GTR = Generalised time-reversible 
    model (\cite{tavare1986some}).
  }
  \label{tab:options}
\bigskip

  \begin{tabular}{|@{}c|p{4cm}|p{9cm}|p{3cm}@{}|}
    \hline
    \centering
    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Description} & \textbf{Includes} \\
    \hline
    $\mathit{G}$ & Generative prior & Diversification model to generate the 'true' tree & $\mathit{t_{p}}$ \\
    $\mathit{A}$ & Alignment model & Model used to simulate the alignment ($\mathit{c_{m}}$ must be strict) & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
    $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 & $\mathit{c_{m}},\mathit{s_{m}},\mathit{t_{p}},\mathit{m_{p}},\mathit{m_{s}}$ \\
    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference & $\mathit{m_{t}}, \mathit{r_{i}}$\\
    $\mathit{X}$ & Experiment & Full setting for the BEAST2 run & $\mathit{I},\mathit{C}$ \\
    $\mathit{E}$ & Error measure parameters & Errors measurement setup & $\mathit{e_{f}}$\\
    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & $\mathit{t_{m}}$\\
    \hline 
  \end{tabular}
  \caption{
    Definitions of terms and relative symbols used in the main text and in Fig~\ref{fig:pipeline}.
  }
  \label{tab:definitions}
  \giovanni{
    Should we rename "Twinning Parameters" to "Twinning model"?
  }
  \richel{
    As the twinning parameters contain more than only the twin model's tree prior,
    I suggest to keep it 'twinning parameters'
  }
  \giovanni{
    Should we rename "twin model" to "twin prior"?
  }
  \richel{
    I think that's a great idea! I would prefer the name 'twin tree prior' even
    better
  }
\end{sidewaystable}

@@

%\input{example_4_esses.latex}
% has label tab:esses_example_4

%\input{example_4_evidences.latex}
% has label tab:evidences_example_4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 5}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
  \centering
  \resizebox {0.8\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.4\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.13\textheight]{example_5_true_alignment.png}
    };   
    \node[state] (CG) [below of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_true_posterior_gen.png}
    };   
    \node[state] (DG) [below of = CG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_error_violin_gen.png}
    };   
    \node[state] (CB) [right of = C, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_true_posterior_best.png}
    };   
    \node[state] (DB) [below of = CB, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_error_violin_best.png}
    };   
    \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
      \includegraphics[height=0.2\textheight]{example_5_twin_tree.png}
    };   
    \node[state] (BT) [below of = AT, rectangle] {
      \includegraphics[height=0.13\textheight]{example_5_twin_alignment.png}
    };   
    \node[state] (CTG) [below of = BT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_gen.png}
    };   
    \node[state] (DTG) [below of = CTG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_gen.png}
    };   
    \node[state] (CTB) [right of = CTG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_best.png}
    };   
    \node[state] (DTB) [below of = CTB, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_best.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south] node {} (CG)
      (CG) edge [anchor = south] node {} (DG)
      (B) edge [anchor = south east] node {} (CB)
      (CB) edge [anchor = south] node {} (DB)
      (A) edge [anchor = east] node {} (AT)
      (AT) edge [anchor = south] node {} (BT)
      (BT) edge [anchor = south east] node {} (CTG)
      (CTG) edge [anchor = south] node {} (DTG)
      (BT) edge [anchor = south] node {} (CTB)
      (CTB) edge [anchor = south] node {} (DTB)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_5_full_pipeline}
  \caption{Example 5: full pipeline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{example_5_esses_gen.latex}
% has label tab:esses_example_5_gen

\input{example_5_esses_best.latex}
% has label tab:esses_example_5_best

\input{example_5_esses_twin_gen.latex}
% has label tab:esses_example_5_twin_gen

\input{example_5_esses_twin_best.latex}
% has label tab:esses_example_5_twin_best

\input{example_5_evidence_true.latex}
% has label tab:evidences_example_5

\input{example_5_evidence_twin.latex}
% has label tab:evidences_example_5

>>>>

\begin{sidewaystable}
\centering
  \begin{tabular}{|@{}c|p{2.5cm}|p{9cm}|p{4.5cm}@{}|}
    \hline
    \centering
    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} & \textbf{Possible values} \\ 
    \hline
    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model & Yule, BD, CBS, CCP, CEP \\
    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates & strict, relaxed log-normal \\
    $\mathit{s_{m}}$ & site model & Nucleotide substitution model & JC, HKY, TN93, GTR \\
    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs & $m_{r} \in \mathbb{R}_{>0}$\\
    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree & any combination of a, c, g, t \\
    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common ancestor & see documentation \\
    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup & see documentation \\
    $\mathit{m_{t}}$ & model type & Criterion to select an inference model & Generative, Candidate \\
    $\mathit{r_{i}}$ & run if & Condition under which an inference model is used & Always, Best candidate\\
    $\mathit{e_{f}}$ & error function & Specifies how to measure the error & nLTT, $|\gamma|$ \\
    $\mathit{t_{m}}$ & twin tree prior & Sets the tree prior for the twin tree & Yule, BD \\
    \hline
  \end{tabular}
  \caption{
    Most important parameter options. BD = birth death, CBS = coalescent
    Bayesian Skyline, CCP = coalescent constant-population, CEP = coalescent
    exponential-population, JC = Jukes-Cantor 1969 (\cite{jukes1969evolution}), HKY = Hasegawa, Kishino and 
    Yano (\cite{hasegawa1985dating}), TN93 = \cite{tamura1993estimation}, GTR = Generalised time-reversible 
    model (\cite{tavare1986some}).
  }
  \label{tab:options}
\bigskip

  \begin{tabular}{|@{}c|p{4cm}|p{9cm}|p{3cm}@{}|}
    \hline
    \centering
    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Description} & \textbf{Includes} \\
    \hline
    $\mathit{P}$ & Generative tree prior & Diversification model to generate the 'true' tree & $\mathit{t_{p}}$ \\
    $\mathit{A}$ & Alignment model & Model used to simulate the alignment ($\mathit{c_{m}}$ must be strict) & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
    $\mathit{G}$ & Generative model & The full setting to produce BEAST2 input data & $\mathit{P}$, $\mathit{A}$ \\
    $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 & $\mathit{c_{m}},\mathit{s_{m}},\mathit{t_{p}},\mathit{m_{p}},\mathit{m_{s}}$ \\
    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference & $\mathit{m_{t}}, \mathit{r_{i}}$\\
    $\mathit{X}$ & Experiment & Full setting for the BEAST2 run & $\mathit{I},\mathit{C}$ \\
    $\mathit{E}$ & Error measure parameters & Errors measurement setup & $\mathit{e_{f}}$\\
    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & $\mathit{t_{m}}$\\
    \hline 
  \end{tabular}
  \caption{
    Definitions of terms and relative symbols used in the main text and in Fig~\ref{fig:pipeline}.
  }
  \label{tab:definitions}
  \giovanni{
    Should we rename "Twinning Parameters" to "Twinning model"?
  }
  \richel{
    As the twinning parameters contain more than only the twin model's tree prior,
    I suggest to keep it 'twinning parameters'
  }
  \giovanni{
    Should we rename "twin model" to "twin prior"?
  }
  \richel{
    I think that's a great idea! I would prefer the name 'twin tree prior' even
    better
  }
\end{sidewaystable}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\iffalse
\begin{figure}[h]
  \centering
  \resizebox {0.8\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.4\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.13\textheight]{example_4_true_alignment.png}
    };   
    \node[state] (CG) [below of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_true_posterior_gen.png}
    };   
    \node[state] (DG) [below of = CG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_error_violin_gen.png}
    };   
    \node[state] (CB) [below right of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_true_posterior_best.png}
    };   
    \node[state] (DB) [below of = CB, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_error_violin_best.png}
    };   
    \node[state] (AT) [right of = A, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_tree.png}
    };   
    \node[state] (BT) [below of = AT, rectangle] {
      \includegraphics[height=0.13\textheight]{example_4_twin_alignment.png}
    };   
    \node[state] (CTG) [below of = BT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_gen.png}
    };   
    \node[state] (DTG) [below of = CTG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_gen.png}
    };   
    \node[state] (CTB) [below right of = BT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_best.png}
    };   
    \node[state] (DTB) [below of = CTB, rectangle] {
      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_best.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south] node {} (CG)
      (CG) edge [anchor = south] node {} (DG)
      (B) edge [anchor = southeast] node {} (CB)
      (CB) edge [anchor = south] node {} (DB)
      (A) edge [anchor = east] node {} (AT)
      (AT) edge [anchor = south] node {} (BT)
      (BT) edge [anchor = southeast] node {} (CTG)
      (CTG) edge [anchor = south] node {} (DTG)
      (BT) edge [anchor = south] node {} (CTB)
      (CTB) edge [anchor = south] node {} (DTB)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_4_full_pipeline}
  \caption{Example 4: full pipeline}
\end{figure}
\fi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{example_4_esses.latex}
% has label tab:esses_example_4

\input{example_4_evidences.latex}
% has label tab:evidences_example_4
